\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{mathrsfs}
\usepackage{svg}
\usepackage{xr}
\usepackage{amsthm,cite,url,graphicx,booktabs,lipsum,color,bm,caption,subcaption,soul}
\usepackage{pifont,tikz,paralist,multirow,amssymb}
\usepackage{xifthen}
\usepackage{enumerate}
\usepackage{titlesec}

% 自定义 Reviewer 计数器
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% 将 Reviewer 的计数方式改为字母（A, B, C, D, ...）
\renewcommand{\thereviewer}{\Alph{reviewer}}

% 定义 Reviewer 的引用格式
\renewcommand{\thepoint}{Comment\,\thereviewer.\arabic{point}}

% 定义 Associate Editor 的评论环境
\newenvironment{point1}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Associate Editor Comment~}} ---\ }
   {\par }

% 定义 Reviewer 的评论环境
\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~\thepoint} } ---\ }
   {\par }

% 定义 Reviewer 的简短评论命令
\newcommand{\shortpoint}[1]{\refstepcounter{point}   \bigskip \noindent 
	{\textbf{Reviewer~Point~\thepoint} } ---~#1  }

% 定义回复环境
\newenvironment{reply}
   {\medskip \noindent \color{blue} \begin{sf}\textbf{Reply}:\  }
   {\medskip \par \end{sf}}

% 定义简短回复命令
\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip \end{sf}}

% 允许跳过某些 Reviewer
\newcommand{\skipreviewer}{\stepcounter{reviewer}}

% 定义 Reviewer 部分
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

% Title
\title{Notes on resubmitting revision made to manuscript No.xxxxx}
\author{}
\date{}

\begin{document}
\maketitle

\vspace{-30pt}

The authors would like to thank the editor and reviewers for their constructive comments and suggestions titled \textit{XX} that have helped improve the quality of this manuscript. The resubmitted manuscript has undergone a thorough revision according to the editor and reviewers’ comments. Please see our responses below. For the reviewers’ convenience, we have highlighted significant changes in the revised manuscript in \textcolor{blue}{blue}.

% Associate Editor 的评论和回复
\section*{Associate Editor Comments}
\begin{point1}
This is a comment from the Associate Editor.
\end{point1}
\begin{reply}
We thank the Associate Editor for this comment. We have addressed this issue in the revised manuscript as follows:
\begin{itemize}
    \item Added a detailed analysis of user study results.
    \item Improved the evaluation of counterfactual explanations for user preferences.
\end{itemize}
\end{reply}

% Reviewer A 的评论和回复
\reviewersection % Reviewer A

\begin{point}
This is a comment from Reviewer A.
\end{point}
\begin{reply}
We thank the reviewer for this comment. We have addressed this issue in the revised manuscript.
\end{reply}

% 跳过 Reviewer B
\skipreviewer % 跳过 Reviewer B

% Reviewer C 的评论和回复
\reviewersection % Reviewer C

\begin{point}
This is a comment from Reviewer C.
\end{point}
\begin{reply}
We appreciate the reviewer's feedback. We have revised the manuscript accordingly.
\end{reply}

% Reviewer D 的评论和回复
\reviewersection % Reviewer D

\begin{point}
This is a comment from Reviewer D.
\end{point}
\begin{reply}
We thank the reviewer for pointing this out. We have clarified this in the revised manuscript.
\end{reply}

\end{document}